\chapter{Joint ISR-Optical Tomography Analysis}
\label{chapter:fusion}
\thispagestyle{myheadings}

\graphicspath{{Fusion/}}

\epigraph{In order to solve the problem of the cause of the aurora and magnetic perturbations, it was necessary to have at our command simultaneous magnetograms and observations from several suitable polar stations at distances of about 1000 kilometers from one another, and also corresponding material from as many other stations all over the world as it was possible to obtain. I demonstrated namely, that certain well-defined magnetic perturbations that occurred over large portions of the earth might be naturally explained as the effect of electric currents.}{\citep{birkeland1908}}

Alfvén waves are an important mechanism for communicating magnetospheric stress into Earth's upper atmosphere.
A significant fraction of the Alfvénic flux is realized as ionospheric Langmuir turbulence and thin dynamic auroral arcs.
Key signatures of Alfvénic dispersion are revealed in the spatiotemporal structure of the aurora. 
Alfvénic auroral structures are strikingly distinct from auroral arcs produced by quasi-static parallel electric fields.
Distinguishing quasi-static arcs from Alfvénic arcs requires high-speed, high-resolution filtered optical systems to capture prompt emissions.
Coordinated optical and incoherent scatter radar measurements routinely show strong echoes from ionospheric destabilization spatiotemporally associated with Alfvénic aurora.
Such observations suggest a connection between Alfvénic particle flux and destabilization of ion-acoustic and Langmuir (electron-acoustic) waves in the lower ionosphere.
The utility of a ground-based high-speed optical tomography system providing estimates of electron precipitation energy spectral characteristics during Alfvén-wave driven destabilization is demonstrated.

\input{./Fusion/intro}
\input{./Fusion/obs}
\input{./Fusion/ana}
\input{./Fusion/disc}